---
title: "Stats Exploration HW 2"
author: "Shalini Mishra"
date: "10/8/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
**Question 1**  

- $H_{0}$ : $\mu \geq 22$
- $H_{a}$ : $\mu < 22$
		

```{r}
#sample size
n <- 220
q1_data <- read.csv('Question1.csv')
head(q1_data)
t.test(q1_data$Payment,alternative="less",mu=22)
#type 1 more serious

```
As p-value=0.1752>0.05, we failed to reject the null hypothesis.
Hence, the CFO can't conclude that the system is profitable.

At 5% significance level, there is not sufficient evidence to support the claim that the proposal of CFO to use SSA (self addressed stamped) envelopes is profitable for the company.

**Question 2**

-  $H_{0}$ : $p_{r} \leq 0.5$
-  $H_{a}$ : $p_{r} > 0.5$
		

```{r}
q2_data <- read.csv("Question2.csv")
#(q2_data)
n <- nrow(q2_data)
num_dem <- length(which(q2_data == 1)) #code 1 for democrat votes
num_rep <- length(which(q2_data == 2)) #code 2 for republican votes
prop.test(num_rep, n, alternative = "greater", p = 0.5, conf.level=0.95)
#reject null hypothesis
```
At 5% significance level, there is sufficient evidence to support the claim that Republicans will win the elections. 
Based on the estimation, the network should announce at 8:01 p.m. that the Republican candidate will win.

**Question 3**  
Let $\mu$ be the weekly spent average for the town.

-  $H_{0}$ : $\mu = 150$
-  $H_{a}$ : $\mu \neq 150$

*a*   
```{r a}
q3_data <- read.csv("Question3.csv")
#head(q3_data)
#At a significance level of 5%
t.test(q3_data$Weekly_food_expense,alternative="two.sided",mu=150)

```

At 5% significance level, p value (say p) =0.000176
- p<0.10 (10% significance) => Reject the null Hypothesis, $H_{0}$
- p<0.05 ( 5% significance) => Reject the null Hypothesis, $H_{0}$
- p<0.01 ( 1% significance) => Reject the null Hypothesis, $H_{0}$

Yes, the sample evidence is statistically significant. We can reject the null hypotheses 
From 1% to 10% significance level, there is sufficient evidence to support the claim that the weekly spent average for the town is different than the national average.

*b*  
```{r b}
#1% significance level
t.test(q3_data$Weekly_food_expense,alternative="two.sided",conf.level=0.99,mu=150)
#Critical values for 1% significance
a <- 150-(qt(0.995,nrow(q3_data)-1)*sd(q3_data$Weekly_food_expense)/(sqrt(nrow(q3_data))))
b <- 150+(qt(0.995,nrow(q3_data)-1)*sd(q3_data$Weekly_food_expense)/(sqrt(nrow(q3_data))))
cat("Any value of sample mean below",a,"and any value above",b,"will reject the null hypothesis at 1% significance level \n")

#10% significance level
t.test(q3_data$Weekly_food_expense,alternative="two.sided",conf.level=0.90,mu=150)
#critical values for 10% significance
a <- 150-(qt(0.95,nrow(q3_data)-1)*sd(q3_data$Weekly_food_expense)/(sqrt(nrow(q3_data))))
b <- 150+(qt(0.95,nrow(q3_data)-1)*sd(q3_data$Weekly_food_expense)/(sqrt(nrow(q3_data))))
cat("Any value of sample mean below",a,"and any value above",b,"will reject the null hypothesis at 10% significance level\n")


```

**Question 4**  
Let $\sigma^2$ be the variance of the daily demand of the product.  

-  $H_{0}$ : $\sigma^2 = 250$
-  $H_{a}$ : $\sigma^2 \neq 250$

```{r}
q4_data <- read.csv("Question4.csv")
#head(q4_data)
n <- nrow(q4_data)
EnvStats::varTest(q4_data$Demand,alternative = "two.sided",sigma.squared=250)
```
p-value=0.708 > 0.05 (5% significance) => Failed to reject H0
The data doesn’t provide enough evidence at the 5% significance level to infer that the operations research analyst’s assumption about the variance is wrong.
At 5% significance level, there is not sufficient evidence to support the claim that variance is different than 250.


**Question 5**  

-  $H_{0}$ : $p_{w} \geq 0.14$
-  $H_{a}$ : $p_{w} < 0.14$

*a.*  
From the alternate hypothesis, we can see it’s a left tailed test.
To determine the sample proportion values that will lead to the show’s cancellation, assuming a 5% significance level,
We will have to find the critical value 
```{r}
z_score <- qnorm(0.95)
p_hat <- 0.14
p_critical <- p_hat - z_score*(sqrt(p_hat*(1-p_hat)/1500))
p_critical
```
the sample proportion values that will lead to the show’s cancellation, assuming a 5% significance level are:
Any sample proportion below 12.5%

Hence ,There is  sufficient evidence at 5% significance level to support the claim that the show have viewership less than 14% to get it cancelled only when sample proportion < 12.5% 

*b*.  
My goal is to find the significance level which will give me the same rejection critical proportion as above

Using the same p-critical value ie. 12.5%, we will try to obtain the corresponding z-score
A sample will reject the null hypothesis if the proportion of people watching it will be below 12.5%
With a new population proportion of 13.4%, we will find the z-score

```{r}
z_score <- (0.125-0.134)/(sqrt((0.134*(1-0.134))/1500))

z_score
pnorm(z_score,mean=0,sd=1,lower.tail=TRUE)

```
The probability that a sample will lead to rejection of the null hypothesis is 15.3% 

**Question 6** 

-  $H_{0}$ : Declare Emergency, fuel=0
-  $H_{a}$ : enough fuel to stay afloat for 15 mins, fuel>0

Error Analysis table is in the word


**Question 7**  

-  $H_{0}$ : the accused is innocent, not enough votes 
-  $H_{a}$ : the accused is guilty, jury votes unanimously

*a.*
According to the question,
Evidence in the trial of an innocent suspect is enough to convince 42% of all jurors in the population that the suspect is guilty.
Each of these 12 jurors are picked from the population.
Probability that a juror is voting for an innocent suspect as guilty =p=0.42
As they are voting unanimously, I am considering each activity of voting as an event

Probability that a jury convicts an innocent suspect 
= p x p x p x…….12 times = p12 = (0.42)12 ~ 0.00003

```{R}
prob_convict_when_innocent <- dbinom(12,12,0.42)
prob_convict_when_innocent
```

*b.*  
Refer to the error analysis table drawn for this question in the word file
Type 1 error was committed by the Jury in part (a)

*c.*  
Evidence in the trial of a guilty suspect is enough to convince 99% of all jurors in the population that the suspect is guilty.
Probability of a juror voting that the suspect is guilty =0.99
Probability of the jury voting that the suspect is guilty(conviction) = 0.99 X 0.99 X …12 times
						            = 0.9912 = 0.88638487 ~ 0.886
Probability of the jury voting that the suspect is not guilty (acquitted) = 1 – Prob(conviction by jury)
							= 1-0.886 = 0.114

```{R}
prob_conv_when_guilty <- dbinom(12,12,0.99)
prob_acquit_when_guilty <- 1- dbinom(12,12,0.99)
prob_acquit_when_guilty

```
*d.*  
From the error analysis table drawn in part(b)
We can conclude that the case in part (c) is a type 2 error.
