---
title: "Stats Exploration Reasoning HW_3"
author: "Shalini Mishra"
date: "10/20/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Question 1**   
Null Hypothesis is that there is no change in productivity between exercisers and non-exercisers  
Alternate Hypothesis is that exercisers are more productive than non-exercisers  
Let, $\mu_{ex}$ be mean productivity for exercisers 
    and $\mu_{noex}$ for non-exercisers  

- $H_{0}$ : $\mu_{ex} -  \mu_{noex} = 0$
- $H_{a}$ : $\mu_{ex} -  \mu_{noex} > 0$

```{r}
exercise1 <- read.csv('Question 1.csv')
#Assessing normality through informal technique
ex1_1 <- na.omit(dplyr::filter(exercise1,Exerciser=='Yes'))
qqnorm(ex1_1$Rating)
qqline(ex1_1$Rating)
ex1_2 <- na.omit(dplyr::filter(exercise1,Exerciser=='No'))
qqnorm(ex1_2$Rating)
qqline(ex1_2$Rating)
```
Because the points are approximately linear, We can conclude that the sample data come from a population that has a normal distribution.  

**a**   
```{r}
library(dplyr)
yes_mean <- exercise1%>%filter(Exerciser=='Yes')%>%
    summarise(num=n(),mean_y=mean(Rating),sd_y=sd(Rating))
no_mean <- exercise1%>%filter(Exerciser=='No')%>%
    summarise(num=n(),mean_n=mean(Rating), sd_n=sd(Rating))
#equal variance test using F test
var.test(ex1_1$Rating,ex1_2$Rating,ratio=1,alternative='two.sided')
#As p value >0.05=> failed to reject null hypothesis: var.equal=TRUE
t.test(ex1_1$Rating,ex1_2$Rating,mu=0, alternative='g',var.equal=TRUE)

```
As p-value<0.05  => Reject null hypothesis,$H_{0}$
At 5% significance level, there is sufficient evidence to support the claim that exercisers outperform non-exercisers on an average i.e. average rating of exercisers are higher than the non-exercisers'.  

**b**    
No, we can't infer that any difference between the two groups is due to exercise. We could have had it been a paired test sample i.e. employees' average rating pre-exercising and post exercising. In the given case, the difference can be attributed to other factors since we are looking at a different groups.

**Question 2**    
Null Hypothesis is that there is no difference between appraisal value and selling prices  
Alternate Hypothesis is that there is a difference  
Let, $\mu_{ap}$ be mean appraisal value
    and $\mu_{sp}$ be mean selling price

- $H_{0}$ : $\mu_{ap} -  \mu_{sp} = 0$
- $H_{a}$ : $\mu_{ap} -  \mu_{sp} \neq 0$
```{r}
exercise2 <- read.csv('Question 2.csv')
head(exercise2)
#F test to compare variances
var.test(exercise2$Value, exercise2$Price, ratio = 1, 
         alternative = "two.sided")
```
Since, p_value<0.05  
At 5% significance level,   
We have sufficient evidence to support the claim that the two variances are not equal:var.equal=FALSE  
```{r}
#As we are considering same 75 cars for both, it's a paired sample
t.test(exercise2$Value, exercise2$Price, alternative='two.sided',
       mu=0,paired=TRUE)
```
Our p-value for the given test is 0.7236 >> 0.05
At 5% significance level, we don't have sufficient evidence to support our claim that difference exists between appraisal value and selling price  
Due to the p-value being very high, For all the below significant values,  
$\alpha$ = 0.01 << 0.7236 => failed to reject $H_{0}$  
$\alpha$ = 0.05 << 0.7236 => failed to reject $H_{0}$  
$\alpha$ = 0.10 << 0.7236 => failed to reject $H_{0}$  
It is appropriate to conclude that no difference exists between appraisal values and selling prices  

**Question 3** 
Null Hypothesis is that there is no difference between variances in service times of teller 1 and 2  
Alternate Hypothesis is that there is a difference  
Let, $\sigma_{1}$ be variance for teller 1 
    and $\sigma_{2}$ be variance for teller 2

- $H_{0}$ : $\sigma_{1} / \sigma_{2} = 1$
- $H_{a}$ : $\sigma_{1} /  \sigma_{2} \neq 1$
```{r}
exercise3 <- read.csv('Question 3.csv')
#head(exercise3)
var.test(exercise3$Teller1,exercise3$Teller2,ratio=1,
         alternative='two.sided', conf.level=0.90)
```
As, p-value<<0.10 => Reject the null hypothesis  
At 10% significance level, we have sufficient evidence to support the claim that variance in service time differ for two tellers.
Yes,The data allow us to infer at the 10% significance level that the variance in service times differs between the two tellers.

**Question 4**  
Null Hypothesis is that prop. of patients who developed serious heart problems using Vioxx and using Placebo are equal
                Or more in placebo group than Vioxx  
Alternate Hypothesis is higher prop. of serious heart patients in Vioxx group than placebo  
Let, $p_{v}$ be prop. of patients who developed serious heart problems using *Vioxx* for 18 months  
and $p_{p}$ be prop. of patients who developed serious heart problems using *placebo* for 18 months  

- $H_{0}$ : $p_{v} -  p_{p} \leq 0$
- $H_{a}$ : $p_{v} -  p_{p} > 0$
For quality assurance and healthcare, we consider significance value of 0.01 (confidence level of 99%).  
```{r}
prop.test(c(45, 25), c(1287, 1299), alternative = "greater", correct = FALSE,
          conf.level=0.99)
```
p-value = 0.006891 < 0.01  
At 1% significance level, we have sufficient evidence to support the claim that the proportion of serious heart patients is higher for those who use Vioxx than placebo.  
```{r}
(45/1287)*100
(25/1299)*100
```
From Vioxx user's point of view,  
Merely 3% of the total Vioxx users and ~2% of the total placebo users suffered from serious heart problems  
Due to very minimal difference between both groups, I as a patient would have attributed this more to chance than something 
which is caused by Vioxx consumption
```{r}
print('Rough/manual Extrapolation for 2M americans')
#serious heart patients for Vioxx
(45/1287) * 2000000
#serious heart patients for non-Vioxx(placebo)
(25/1299) * 2000000

print('Using confidence intervals')
### For 99% confidence level, find the chi-sq (lower) and chi-sq (upper)
#z score for 95% 
z99 <- qnorm(0.995,0,1,TRUE)
#proportion of heart patients for Vioxx
p_v <- 45/1287
#upper confidence limit
ucl_v <- p_v+ z99*sqrt((p_v*(1-p_v))/1287)
#lower confidence limit
lcl_v <- p_v- z99*sqrt((p_v*(1-p_v))/1287)
cat("99 percent Confidence Interval for Vioxx Heart patients proportion in true population:\n", round(lcl_v,3), round(ucl_v,3))

round(lcl_v,3)*2000000
round(ucl_v,3)*2000000
#proportion of heart patients for Placebo
p_p <- 25/1299
#upper confidence limit
ucl_p <- p_p+ z99*sqrt((p_p*(1-p_p))/1299)
#lower confidence limit
lcl_p <- p_p- z99*sqrt((p_p*(1-p_p))/1299)
cat("99 percent Confidence Interval for placebo Heart patients proportion in true population:\n", round(lcl_p,3), round(ucl_p,3))

round(lcl_p,3)*2000000
round(ucl_p,3)*2000000
```
Analyzing simple manual extrapolation:
69K Vioxx users with serious health problems compared to 38K non vioxx users is a result which isn't something Merck would 
like to pursue and moreover they're inclined to avoid it. 
Analyzing extrapolation using CI:  
Roughly between 44K and 96K people can contain the number of true population heart patients.it still is staggeringly high compared to the estimates for Placebo. That's a risk any pharma company would like to avoid  
The results are both statistically significant(test result) and practically significant for Merck as a company to close its production and pull it out from the market   
Merck can experience following losses:  
Hamper their reputation which have both long term and short term repurcussions  
Can instigate legal battle due to striking difference in numbers -expensive  
Brach of trust and hampering of human lives     

**Question 5**  

**a**.   
125 employees from West Coast - increased health benefits  
240 employees from East Coast - increased vacation days  
This is a case of independant samples. There can be many confounding factors  
As per my understanding, confounding effect is something which hasn't been accounted for and can certainly hamper the
final result of an experiment. In our experiment, we are considering samples from two different places. There can
be confounding due to:  
-- Difference in location (east/west coast)
-- Distribution of employess as per Gender
-- Age Group - Irregular distribution of age groups in these samples can give inconsistent results
-- Departments under the company considered for each sample in this analysis- nature of retention might be different
   across different departments
-- Difference in sample size

**b**   
Null Hypothesis is that difference in prop. of employees retained between groups with increased health benefits and increased vacation days is only 0.05 or less   
Alternate Hypothesis is higher prop. of employees retained from health benefits group by 0.05 or more  
Let, $p_{h}$ be prop. of employees retained from the group with increased health benefits  
and $p_{v}$ be prop. of employees reatined from the group with increased vacation days  

- $H_{0}$ : $p_{h} -  p_{v} \geq 0.05$
- $H_{a}$ : $p_{h} -  p_{v} < 0.05$
```{r}
exercise5 <- read.csv('Question 5.csv')
#head(exercise5)
#nrow(exercise5)
#prop. of employees retained from the group with increased health benefits
ret_h <- length(which(exercise5$Benefit == 'Health' & exercise5$Retention == 1))
n_h <- 125
prop_h <- ret_h/n_h
#prop. of employees retained from the group with increased vacation days 
ret_v <- length(which(exercise5$Benefit == 'Vacation' & exercise5$Retention == 1))
n_v <- 140
prop_v <- ret_v/n_v
#appropriate test statistic
z_score <- ((prop_h-prop_v) - 0.05)/sqrt((prop_h*(1-prop_h)/n_h)+(prop_v*(1-prop_v)/n_v))

p_value <- pnorm(z_score)
p_value
```

Since,  p_Value >  $\alpha$ = 0.01 => Failed to reject $H_{0}$  
        p_Value >  $\alpha$ = 0.05 => Failed to reject $H_{0}$  
        p_Value >  $\alpha$ = 0.10 => Failed to reject $H_{0}$  
At all the significance levels mentioned above, there is not sufficient evidence to warrant rejection of the the claim that the difference in retention proportion will be higher than or equal to 0.05.  
Yes,we have statistically significant evidence to support the fact we have higher retention to compensate for switching to health benefits.  

**c**    
Null Hypothesis is that there is no difference in prop. of employees retained between groups with increased health benefits and increased vacation days  
Alternate Hypothesis is there is difference between the retention proportion of the two groups  
Let, $p_{h}$ be prop. of employees retained from the group with increased health benefits  
and $p_{v}$ be prop. of employees reatined from the group with increased vacation days  

- $H_{0}$ : $p_{h} -  p_{v} = 0$
- $H_{a}$ : $p_{h} -  p_{v} \neq 0$
```{r}
prop.test(c(ret_h, ret_v),c(n_h,n_v), alternative = "two.sided",
          correct = FALSE)
```
p_value=0.1051 > 0.05 => Failed to reject $H_{0}$  
p_value=0.1051 > 0.10 => Failed to reject $H_{0}$  
At 5% significance level, there is not sufficient evidence to support the claim that there is a difference in retention proportions for the two groups.  
Same is the case for 10% significance level  
No, there is no statistically significant difference in retention rates between the benefit plans based on the evidence provided to us.  

**Question 6**
```{r}
#Reported incomes for 2000,2008,2014
exercise6 <- read.csv('Question 6.csv')
#CPI inndices for each year
annual_data <- readxl::read_excel('U.S. CPI Annual.xlsx')

head(exercise6)
nrow(exercise6)
head(annual_data)
colnames(annual_data) <- c('Year','CPI')
```

The Consumer Price Index (CPI) is a measure of the average change over time in the prices paid by urban consumers for a market basket of consumer goods and services. Indexes are available for the U.S. and various geographic areas. Average price data for select utility, automotive fuel, and food items are also available. (Source-https://www.bls.gov/cpi/)

**a**   
Let, $\mu_{2008}$ be mean income in 2008  
    and $\mu_{2000}$ be mean income in 2000  

- $H_{0}$ : $\mu_{2008} -  \mu_{2000} \leq 0$
- $H_{a}$ : $\mu_{2008} -  \mu_{2000} > 0$ i.e.$\mu_{2008}$ > $\mu_{2004}$

Please note we are not adjusting for inflation here
```{r}
income_2000 <- na.omit(dplyr::select(exercise6,RINCOME_2000))
income_2008 <- na.omit(dplyr::select(exercise6,RINCOME_2008))
#Conducting a test to check if population variances are equal using F test
var.test(income_2008$RINCOME_2008,income_2000$RINCOME_2000,ratio=1,
         alternative = "two.sided")
#As p value <0.05, we reject the null hypothesis i.e. var.equal=FALSE

#As these are not same set of people, these are independant samples(paired=FALSE)
t.test(income_2008,income_2000,alternative='greater',mu=0,
       paired=FALSE)
```
As, p-value << 0.05 => Reject the null hypothesis  
At 5% significance level,We have sufficient evidence to support the claim that income rose between 2000 and 2008 

**b**   
Let, $\mu_{2014}$ be mean income in 2014
    and $\mu_{2008}$ be mean income in 2008

- $H_{0}$ : $\mu_{2014} -  \mu_{2008} \leq 0$
- $H_{a}$ : $\mu_{2014} -  \mu_{2008} > 0$

Please note we are not adjusting for inflation here
```{r}
income_2008 <- na.omit(dplyr::select(exercise6,RINCOME_2008))
income_2014 <- na.omit(dplyr::select(exercise6,RINCOME_2014))
#Conducting a test to check if population variances are equal using F test
var.test(income_2014$RINCOME_2014,income_2008$RINCOME_2008,ratio=1,
         alternative = "two.sided")
#As p value <0.05, we reject the null hypothesis i.e. var.equal=FALSE

#As these are not same set of people, these are independant samples(paired=FALSE)
t.test(income_2014,income_2008,alternative='greater',mu=0,
       paired=FALSE)
```
As, p-value < 0.05 => Reject the null hypothesis  
At 5% significance level,We have sufficient evidence to support the claim that income rose between 2008 and 2014  

**c**  
Let, $\mu_{adj}$ be inflation adjusted mean income in 2008 (w.r.t 2000)
    and $\mu_{2000}$ be mean reported income in 2000

- $H_{0}$ : $\mu_{adj} -  \mu_{2000} \leq 0$
- $H_{a}$ : $\mu_{adj} -  \mu_{2000} > 0$
```{r}
library(dplyr)
CPI_08 <- annual_data%>% select(CPI)%>% filter(annual_data$Year==2008)
CPI_00 <- annual_data%>% select(CPI)%>% filter(annual_data$Year==2000)
inf_rate <- (CPI_08-CPI_00)/CPI_00
#inflation rate in 2008 considering 2000 as the base year
inf_rate
```

```{r}
class(income_2008)
adj_factor <- rep((1-inf_rate),length(income_2008))
income_2008 <- cbind(income_2008,adj_factor)
#head(income_2008)
income_2008$adjusted_income <- income_2008$RINCOME_2008*income_2008$CPI
#Conducting a test to check if population variances are equal using F test
var.test(income_2008$adjusted_income,income_2000$RINCOME_2000,ratio=1,
         alternative = "two.sided")
#Reject null hypothesis i.e. var.equal=FALSE

t.test(income_2008$adjusted_income,income_2000,alternative='greater',mu=0,
       paired=FALSE)

```
As p-value>> 0.05 => Failed to reject the null hypothesis,  
At 5% significance level, We don't have sufficient evidence to support the claim that income rose between 2000 to 2008 i.e. average income in 2008 was higher than the average income in 2000  

**d**  
Let, $\mu_{adj}$ be inflation adjusted mean income in 2014 (w.r.t 2008)
    and $\mu_{2008}$ be mean reported income in 2008

- $H_{0}$ : $\mu_{adj} -  \mu_{2008} \leq 0$
- $H_{a}$ : $\mu_{adj} -  \mu_{2008} > 0$
```{r}
library(dplyr)
#obtaining CPI index for 2008 and 2014
CPI_08 <- annual_data%>% select(CPI)%>% filter(annual_data$Year==2008)
CPI_14 <- annual_data%>% select(CPI)%>% filter(annual_data$Year==2014)
inf_rate <- (CPI_14-CPI_08)/CPI_08
#inflation rate in 2014 considering 2008 as the base year
inf_rate
```

```{r}
#Real income in 2014 (compared to 2008)=(Reported Income in 2014)*(1- Inflation rate)
adj_factor <- rep((1-inf_rate),length(income_2014))
income_2014 <- cbind(income_2014,adj_factor)
head(income_2014)
income_2014$adjusted_income <- income_2014$RINCOME_2014*income_2014$CPI

#Conducting a test to check if population variances are equal using F test
var.test(income_2014$adjusted_income,income_2008$RINCOME_2008,ratio=1,
         alternative = "two.sided")
#failed to reject the null hypothesis, var.equal=TRUE
t.test(income_2014$adjusted_income,income_2008$RINCOME_2008,alternative='greater',mu=0,
       paired=FALSE,var.equal=TRUE)
```
As p-value>> 0.05 => Failed to reject the null hypothesis,  
At 5% significance level, We don't have sufficient evidence to support the claim that income rose between 2008 to 2014 i.e. average income in 2014 was higher than the average income in 2000

**e**    
I have learned that just comparing raw incomes of two different years is an incorrect way. We have to treat for inflation to have a fair omparison between the years. Otherwise there is high possibility of us making biased decisions which in turn will affect the business decisions we make.
we alsways have to make sure when we are comapring, the base should be same